---
title: "UKBB SNP-Heritability Browser"
date: "Last updated `r format(Sys.Date())`"
output: html_document
params:
  isUnivar: FALSE
  datfile: "../results/ukbb_h2part.RData"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# parameters
load(file=normalizePath(params$datfile))
isUnivar <- as.logical(params$isUnivar)

# packages
require(reshape2)
require(ggplot2)
require(plotly)
require(DT)
require(crosstalk)
require(crosstool)

# define utility columns
dat$isBinary <- !is.na(dat$N_case)
dat$Neff <- dat$N
dat$Neff[dat$isBinary] <- round( (4/((1/dat$N_case)+(1/dat$N_control)))[dat$isBinary], 2)
#dat$isNomSig_h2 <- dat$h2_p < .05
#dat$isBonfSig_h2 <- dat$h2_p < (.05/nrow(dat))
#dat$isPassing <- (dat$N > 10000) & (!dat$isBinary | dat$N_case > 400)
#coef_p_cols <- names(dat)[grep("..Coefficient_p",names(dat))]

# setup links
dat$table_desc <- paste0("<a href='h2_summary_",dat$phenotype,".html'>",dat$description,"</a>")

```

<br>

`r if(isUnivar){paste0("<div class=\"panel panel-warning\"><div class=\"panel-heading\"><h3 class=\"panel-title\">Note</h3></div><div class=\"panel-body\">You are viewing results from univariate LD score regression. We recommend the <a href=\"h2_browser.html\">partitioned results</a> instead.</div></div>")}`

```{r lookup, echo=FALSE}
d2 <- dat[,c("phenotype","table_desc","N","prevelence","intercept","intercept_p","h2_liability","h2_p")]
shared_dat <- SharedData$new(d2)

datatable(shared_dat, 
		  rownames = F, 
		  colnames = c("ID","Phenotype","N","Prev.","Int.","Int. p","h2","h2 p"), 
#		  extensions='FixedHeader', 
		  selection="none",
		  style="bootstrap", 
		  class="display", 
		  escape=F,
		  options = list(scrollY="400px", scrollX=TRUE, pageLength=10) # fixedHeader=TRUE, 
) %>% formatSignif(columns = c(5), digits=4) %>% formatSignif(columns = c(6:8), digits=3) %>% formatSignif(columns = 4, digits=2) %>% formatRound(columns=3, digits=0)
```

<br>

<div class="well">

### Lookup a phenotype

Explore results of `r ifelse(isUnivar, "univariate", "partitioned")` LD score regression for UK Biobank phenotypes. You can read more about this analysis and find links to more information [on the annoucement page](http://www.nealelab.is/blog/2017/9/15/heritability-of-2000-traits-and-disorders-in-the-uk-biobank).

You can search for phenotypes by:

* **Description**, e.g. "height"
* **ICD code**, e.g. "I25"
* **ID**, i.e. the UK Biobank phenotype ID matching the [UKBB showcase](http://biobank.ctsu.ox.ac.uk/crystal/search.cgi)

Results shown here include:

* **N**: The sample size in the UKBB analysis set (more info on the GWAS analysis [here](http://www.nealelab.is/blog/2017/9/11/details-and-considerations-of-the-uk-biobank-gwas))
* **Prevalence** (Prev.) for binary phenotypes
* **Intercept** (Int.): the intercept term from `r ifelse(isUnivar, "univariate", "partitioned")` LD score regression. Values > 1 may indicate population stratification, confounding, or some other model misspecification. P-value (Int. p) corresponds to the test that the intercept is greater than 1.
* **Heritability** ($h^2$): the SNP heritability estimate from `r ifelse(isUnivar, "univariate", "partitioned")` LD score regression. Value is given on the liability scale for binary phenotypes, assuming that the population prevalence matches the prevalence in the UK Biobank analysis set. P-value (h2 p) corresponds to the test that the intercept is greater than 1.

Click on a phenotype for more detailed results from the partitioned LD score regression analysis.

</div>
