---
title: "Details on LD Score Regression in UK Biobank"
date: "Last updated `r format(Sys.Date())`"
author: "Results from the [Neale Lab](credits.html)"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: false
    toc_depth: 3
---


```{r child = '_toc_fix.Rmd'}
```

```{r child = '_code_highlight_fix.Rmd'}
```

<br>

<div class="well">

### Methods Outline

* Genotyping and phenotyping of ~500,000 individuals performed by the [UK Biobank](https://www.ukbiobank.ac.uk/). Accessed under application 31063.

* Phenotypes processed for GWAS using [PHESANT](https://github.com/astheeggeggs/PHESANT) and by extracting ICD10 codes (letter + digits) and recreating [FinnGen](https://www.finngen.fi/en) endpoints.

* GWAS performed by the [Neale Lab](https://www.nealelab.is/uk-biobank) using [Hail](https://hail.is) on the [Google Cloud Platform](https://cloud.google.com/). [A full technical discussion](http://www.nealelab.is/blog/2017/9/14/heritability-501-ldsr-based-h2-in-ukbb-for-the-technically-minded) on the limitations of this analysis and recommendations for intepreting the results based on the Round 1 results, as well as more general details about [our analysis of UK Biobank](http://www.nealelab.is/blog/2017/9/11/details-and-considerations-of-the-uk-biobank-gwas), are available elsewhere. Code is also available [on Github](https://github.com/Nealelab/UK_Biobank_GWAS).

* LD Score regression (LDSR; [Bulik-Sullivan et al. 2015](http://www.nature.com/ng/journal/v47/n3/full/ng.3211.html) or [on pubmed](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4495769/)) formatted sumstats for HapMap3 sites passing QC (MAF > 0.01, INFO > 0.9, excluding multiallelics and the MHC region) in UK Biobank extracted with [Hail](https://hail.is).

* SNP Heritability estimated using partitioned LDSR run using the [MTAG](https://github.com/omeed-maghzian/mtag) implementation of [ldsc](https://github.com/bulik/ldsc) with the Baseline-LD v1.1 model ([Gazal et al. 2017](http://www.nature.com/ng/journal/vaop/ncurrent/full/ng.3954.html)) on the [Google Cloud Platform](https://cloud.google.com/).

* Restricted primary SNP heritability analysis to [one GWAS per phenotype](select_topline.html). This includes identifying sex-specific phenotypes, choice of normalization for continuous phenotypes, evaluation of the dilution fraction for biomarker phenotypes, and pruning of [FinnGen](https://www.finngen.fi/en) endpoints that are identical within current UKB data as detailed [here](select_topline.html).

* Results for each primary GWAS evaulated for [confidence](confidence.html) based on power at potential biases as a function of sample size, the expected magnitude of the standard error on the SNP heritability estimate, and potential limitations to interpretation based on sex differences or ordinal encodings in phenotyping.

* Among results with acceptable [confidence](confidence.html), tested [statistical significance](significance.html) of the LDSR SNP heritability estimate accounting for the effective multiple testing burden. More detail on the evaluation of potential thresholds for statistical significance in this analysis [here](significance.html).

* Partitioning results from LDSR reported for phenotypes with strong SNP heritability results ($z > 7$). See phenome-wide [plots](plots_home.html) and access per-phenotype results from the [SNP heritability browser](h2_browser.html).

* Full code for the LDSR analysis is available [on Github](https://github.com/Nealelab/UKBB_ldsc_scripts)

</div>

<br>

<div class="well">

### Available Downloads

Results files for this LDSR analysis are available <a href="https://www.dropbox.com/sh/g28ee03vy4hvqw5/AADAkDbSFHsYE8bme1Jjeekca/ldsc_results?dl=0" onclick="trackOutboundLink('https://www.dropbox.com/sh/g28ee03vy4hvqw5/AADAkDbSFHsYE8bme1Jjeekca/ldsc_results?dl=0'); return false;">on Dropbox</a>. Available files are:

* `ukb31063_h2_topline.02Oct2019.tsv.gz`: Primary SNP heritability results for each phenotype, as reported in the [browser](h2_browser.html).
* `ukb31063_h2_all.02Oct2019.tsv.gz`: LDSR results for all of the UKB Neale Lab GWAS, including partitioning, sex-stratified results, and alternative versions of phenotypes (e.g. biomarker dilution, non-normalized).

In addition to the LDSR results, the exported sumstat files in LDSC format are also available [on Dropbox](). A manifest for easier programatic access is available [here]().

LD scores, weights, and allele frequencies for the Baseline-LD v1.1 model used here are available for [download from the Alkes Price lab](https://data.broadinstitute.org/alkesgroup/LDSCORE/).

Lastly, the previous results based on the Round 1 Neale Lab GWAS remain available for download <a href="https://www.dropbox.com/sh/g28ee03vy4hvqw5/AADAkDbSFHsYE8bme1Jjeekca/ldsc_results?dl=0" onclick="trackOutboundLink('https://www.dropbox.com/sh/g28ee03vy4hvqw5/AADAkDbSFHsYE8bme1Jjeekca/ldsc_results?dl=0'); return false;">on Dropbox</a>, including both univariate and paritioned LDSR results. Round 1 LDSR results can be browsed [here](round1_h2_browser.html) (with the corresponding univariate results [here](round1_h2_univar_browser.html)).

</div>

<br>

<div class="well">

### Additional LDSR Details

* Sumstats files were extracted directly from the UKB results files rather than running `munge_sumstats.py` from [ldsc](https://github.com/bulik/ldsc). This includes directly extracting HapMap3 sites rather than using the precomputed `w_hm3.snplist` reference file.

* LDSR was run using the implementation in [MTAG](https://github.com/omeed-maghzian/mtag) because it provides a convenient interface to [ldsc](https://github.com/bulik/ldsc) from within python rather than via the command line.

* The partitioned LDSR analysis was run using the v1.1 of the Baseline-LD annotations described by [Gazal et al. 2017](http://www.nature.com/ng/journal/vaop/ncurrent/full/ng.3954.html) (also on [biorxiv](http://www.biorxiv.org/content/early/2017/04/03/082024)) computed from 1000 Genomes Phase 3 data from European ancestry populations and corresponding 1000 Genomes allele frequencies (available from [the ldsc reference downloads page](https://data.broadinstitute.org/alkesgroup/LDSCORE/)). 

* Default settings were used with the exception of removing the maximum $\chi^2$ filter in order to avoid bias in phenotypes that have extremely large $\chi^2$ for loci with strong effect sizes in the large sample size of UKB.

* We previously ran both univariate and partitioned heritability analyses. For the reasons written about [here](http://www.nealelab.is/blog/2017/9/20/insights-from-estimates-of-snp-heritability-for-2000-traits-and-disorders-in-uk-biobank), we’re currently presenting the partitioned heritability LDSR results as the primary analysis in the [summary browser](h2_browser.html).

</div>

<br>

<div class="well">

### Site Design

* This site has been generated using [R Markdown](http://rmarkdown.rstudio.com/).

* Interactive plots generated with with [plotly](https://plot.ly/r/) and [crosstalk](https://rstudio.github.io/crosstalk/).

* Searchable tables generated using [DT](https://rstudio.github.io/DT/).

* CSS theme based on [Bootswatch](https://bootswatch.com/3/) themes Flatly and Sandstone, modified with the assistance of [Bootstrap Live Customizer](https://www.bootstrap-live-customizer.com/).

* Favicon generated with the assistance of [Favic-o-matic](http://www.favicomatic.com/).

* Site hosted via [Github Pages](https://pages.github.com/).

* All code for generating the site available on [Github](https://github.com/nealelab/ukbb_ldsc).

</div>

<br>

<div class="well">

### Contact

If you want more information, please **contact us at: <nealelab.ukb@gmail.com>.**

</div>
