---
title: "Comparing Round 1 and Round 2 LDSR Results"
date: "Last updated `r format(Sys.Date())`"
author: "Results from the [Neale Lab](credits.html)"
output: 
  html_document:
    toc: true
    toc_float: true
params:
  datfile1: "../results/round1/ukbb_h2part.RData"
  datfile2: "../results/round2_final/ukb31063_h2_topline.02Oct2019.tsv.gz"
---

```{r child = '_toc_fix.Rmd'}
```

```{r smallp_func, include=FALSE}

# handle intercept_p as a string (due to high precision p-vals)
pstring <- function(p, prec=3){

intp_string <- as.character(p)
intp_string[p == as.character(as.numeric(p))] <- as.character(signif(as.numeric(p),prec)[(p == as.character(as.numeric(p)))])
highstr <- intp_string[!(p == as.character(as.numeric(p)))]
highstr_lead <- as.character(signif(as.numeric(substr(highstr,1,6)),prec))
highstr_exp <- sapply(highstr,function(a) strsplit(a,split="e")[[1]][2],USE.NAMES = F)
intp_string[!(p == as.character(as.numeric(p)))] <- paste0(highstr_lead,"e",highstr_exp)
while(any(intp_string=="1")){
  prec <- prec+1
  intp_string[intp_string=="1"] <- as.character(signif(as.numeric(p),prec)[intp_string=="1"])
}
return(intp_string)
}


```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(reshape2)
require(ggplot2)
require(plotly)
require(DT)
require(crosstalk)
require(crosstool)
require(Rmpfr)
require(htmlwidgets)

# handle params
load(file=normalizePath(params$datfile1))
dat1 <- dat[,c("phenotype","N","N_case","N_control","prevelence","mean_chi2","lambdaGC","intercept","intercept_se","intercept_z","intercept_p","ratio","h2_liability","h2_liability_se","h2_z","h2_p")]
rm(dat)
names(dat1) <- gsub("intercept","round1_intercept",names(dat1))
names(dat1) <- gsub("h2","round1_h2",names(dat1))
names(dat1) <- gsub("ratio","round1_ratio",names(dat1))

# fix tiny int p-value
dat1$round1_intercept_p <- format(mpfr(2,64)*pnorm(mpfr(dat1$round1_intercept_z,64),lower=F), max.digits=15, scientific=T)
dat1$round1_intercept_p <- pstring(dat1$round1_intercept_p)



dat2 <- read.delim(file=normalizePath(params$datfile2),sep='\t',header=T,stringsAsFactors=F,colClasses=c(intercept_p="character"))
dat2 <- dat2[,c("phenotype","description","intercept","intercept_se","intercept_z","intercept_p","ratio","h2_liability","h2_liability_se","h2_z","h2_p","confidence","n","Neff","n_cases","n_controls","isBinary","variable_type","source")]
names(dat2) <- gsub("intercept","round2_intercept",names(dat2))
names(dat2) <- gsub("h2","round2_h2",names(dat2))
names(dat2) <- gsub("ratio","round2_ratio",names(dat2))
dat2$round2_intercept_p <- pstring(dat2$round2_intercept_p)

# specify irnt where needed
# since round 1 continuous vars were all irnt, but doesn't apply to ordinal phens
d1irnt <- paste0(dat1$phenotype,"_irnt")
d1irnt_idx <- (!(dat1$phenotype %in% dat2$phenotype)) & (d1irnt %in% dat2$phenotype)
dat1$phenotype[d1irnt_idx] <- d1irnt[d1irnt_idx]

dat <- merge(dat1,dat2,by="phenotype")

# setup extra fields
dat$r2_int_nlogp <- as.numeric(-log10(mpfr(0.5,64)*erfc(mpfr(dat$round2_intercept_z,64)/sqrt(mpfr(2,64)))))
dat$r1_int_nlogp <- as.numeric(-log10(mpfr(0.5,64)*erfc(mpfr(dat$round1_intercept_z,64)/sqrt(mpfr(2,64)))))

# for clickable data points
dat$description <- paste0("[",dat$phenotype,"] ",dat$description)
dat$urls <- paste0("h2_summary_",dat$phenotype,".html")

# https://stackoverflow.com/questions/56084819/open-link-on-datapoint-click-with-plotly-in-r-shiny 
js <- "
function(el, x) {
  el.on('plotly_click', function(d) {
    if(d.event.button === 0){
      var point = d.points[0];
      var url = point.data.customdata[point.pointIndex];
      window.location.href = url;
    }
  })
}"
```

<br>

### Overview

<div class="well">

Comparison of our current primary results from partitioned LDSR of the [Round 2 Neale Lab GWAS](http://www.nealelab.is/uk-biobank) to previous results from paritioned LDSR of [Round 1 GWAS](http://www.nealelab.is/blog/2017/7/19/rapid-gwas-of-thousands-of-phenotypes-for-337000-samples-in-the-uk-biobank). Both results use the Baseline LD v1.1 model. The Round 2 GWAS results use a slightly broader sample set than 'white British' and incorporate additional covariates ([more detail on github](https://github.com/Nealelab/UK_Biobank_GWAS)).

</div>

<br>

### Intercept {.tabset .tabset-fade}

Since the sample sizes are different, we focus here on comparison of the ratio. Smaller ratios in the current results from Round 2 GWAS suggest reduced confounding, likely due to the expanded set of covariates. P-values for the intercept estimates are similarly less significant in the Round 2 results.

#### LDSR Ratios

<div class="well">
  
```{r int_ratio, echo=FALSE}
shared_int2 <- SharedData$new(dat[!is.na(dat$round1_ratio) & !is.na(dat$round2_ratio) & dat$confidence != "none",])

pp <- plot_ly(shared_int2,
              x = ~round1_ratio,
              y = ~round2_ratio,
              type="scatter",
              mode="markers",
              customdata=~urls,
              hoverinfo="text",
              text = ~paste0(
                "Phenotype: ", description,
                "<br>Round 1 Ratio: ", round(round1_ratio,3), " (Int.=",round(round1_intercept,4), ", p=",round1_intercept_p,")",
              "<br>Round 2 Ratio: ", round(round2_ratio,3), " (Int.=",round(round2_intercept,4), ", p=",round2_intercept_p,")",
              "<br>Round 2 Effective N: ", Neff)
            ) %>% add_trace(
                x=~round1_ratio,
                y=~round1_ratio,
                mode="lines",
                hoverinfo="none",
                showlegend=F
            ) %>% onRender(js)

# construct
bscols(widths=c(10,2),
       config(pp, collaborate = F, showLink=F, displayModeBar="hover", displaylogo=F, sendData=F),
       list(
      		filter_slider("Neff", "Effective N", shared_int2, ~Neff, step=100, width="100%",round=T),
      		filter_select("variable_type","Var. Type", shared_int2, ~variable_type, multiple=T),
      		filter_select("confidence","Confidence", shared_int2, ~confidence, multiple=T),
      		filter_select("source","Dataset", shared_int2, ~source, multiple=T)
       )
)

```

<p></p>

*Note:* Comparison limited to phenotypes with at least some [confidence](confidence.html) in the Round 2 results. Trend is clearer when filtering to higher effective sample size using the slider on the right.

</div>

<br>

#### Intercept estimates
  
<div class="well">
  
```{r comp_int, echo=FALSE}
# setup shared data
shared_int <- SharedData$new(dat[dat$confidence != "none",])

# define plot
pp <- plot_ly(shared_int, 
              x=~round1_intercept, 
              y=~round2_intercept, 
              type="scatter",
              mode="markers",
              customdata=~urls,
              hoverinfo="text",
              text = ~paste0(
                "Phenotype: ", description,
                "<br>Round 1 Int.: ", round(round1_intercept,5), " (p=",round1_intercept_p,")",
                "<br>Round 2 Int.: ", round(round2_intercept,5), " (p=",round2_intercept_p,")",
                "<br>Round 2 Effective N: ", Neff)
            ) %>% add_trace(
                x=~round1_intercept,
                y=~round1_intercept,
                mode="lines",
                hoverinfo="none",
                showlegend=F
            ) %>% onRender(js)

# construct
bscols(widths=c(10,2),
       config(pp, collaborate = F, showLink=F, displayModeBar="hover", displaylogo=F, sendData=F),
       list(
      		filter_slider("Neff", "Effective N", shared_int, ~Neff, step=100, width="100%",round=T),
      		filter_select("variable_type","Var. Type", shared_int, ~variable_type, multiple=T),
      		filter_select("confidence","Confidence", shared_int, ~confidence, multiple=T),
      		filter_select("source","Dataset", shared_int, ~source, multiple=T)
       )
)
```

<p></p>

*Note:* Comparison limited to phenotypes with at least some [confidence](confidence.html) in the Round 2 results.

</div>

<br>
  
#### Intercept p-values
  
<div class="well">
  
```{r qq_int, echo=FALSE}

shared_int3 <- SharedData$new(dat[dat$confidence != "none",])

exp_nlogp <- function(p){
  nn <- length(p)
  qquad <- 1:nn
  qref <- (qquad-.5)/nn
  idx <- rank(p)
  return(-log10(qref[idx]))
}

nnr <- nrow(dat[dat$confidence != "none",])
qquad <-c(1:nnr)
qref <- ((qquad-.5)/nnr)
ci_up <- qbeta(.975,qquad,nnr+1-qquad)
ci_lo <- qbeta(.025,qquad,nnr+1-qquad)

pp <- plot_ly(shared_int3, 
              x=~exp_nlogp(-round1_intercept_z), 
              y=~r1_int_nlogp,
              type="scatter",
              mode="markers",
              customdata=~urls,
              showlegend=T,
              name="Round 1",
              hoverinfo="text",
              text = ~paste0(
                "Phenotype: ", description,
                "<br>Round 1 Int.: ", round(round1_intercept,5), " (p=",round1_intercept_p,")",
                "<br>Round 2 Int.: ", round(round2_intercept,5), " (p=",round2_intercept_p,")",
                "<br>Round 2 Effective N: ", Neff)
          ) %>% add_trace(
              x=~exp_nlogp(-round2_intercept_z), 
              y=~r2_int_nlogp,
              type="scatter",
              mode="markers",
              customdata=~urls,
              showlegend=T,
              name="Round 2",
              hoverinfo="text",
              text = ~paste0(
                "Phenotype: ", description,
                "<br>Round 1 Int.: ", round(round1_intercept,5), " (p=",round1_intercept_p,")",
                "<br>Round 2 Int.: ", round(round2_intercept,5), " (p=",round2_intercept_p,")",
                "<br>Round 2 Effective N: ", Neff)
          ) %>% add_trace(
              x=-log10(qref),
              y=-log10(qref),
              mode="lines",
              showlegend=F,
              hoverinfo = "text",
              text = ""
          ) %>% add_trace(
              x=-log10(qref),
              y=-log10(ci_up),
              mode="lines",
              line=list(color='#2ca02c'),
              showlegend=F,
              hoverinfo = "text",
              text = ""
          ) %>% add_trace(
              x=-log10(qref),
              y=-log10(ci_lo),
              mode="lines",
              line=list(color='#2ca02c'),
              fill="tonexty",
              fillcolor='rgba(44,160,44,0.2)',
              showlegend=F,
              hoverinfo = "text",
              text = ""
          ) %>% layout(
              showlegend=T,
              xaxis = list(title="Expected -log10(p-value)"),
              yaxis = list(title="Observed -log10(p-value)")
          ) %>% onRender(js)

bscols(widths=c(12),
       config(pp, collaborate = F, showLink=F, displayModeBar="hover", displaylogo=F, sendData=F)
)
```

<p></p>

*Note:* Comparison limited to phenotypes with at least some [confidence](confidence.html) in the Round 2 results.

</div>

<br>

### SNP Heritability {.tabset .tabset-fade}

SNP heritability results are highly concordant between Round 1 and Round 2. SNP heritability is increased for a handful of phenotypes where the updated GWAS covariates are likely to have strong effects. P-values for the Round 2 SNP heritability results are slightly stronger in the increased sample size.
  
#### SNP Heritability Estimates
  
<div class="well">
  
```{r comp_h2, echo=FALSE}
# setup shared data
shared_h2 <- SharedData$new(dat[dat$confidence != "none",])

# define plot
pp <- plot_ly(shared_h2, 
              x=~round1_h2_liability, 
              y=~round2_h2_liability, 
              type="scatter",
              mode="markers",
              customdata=~urls,
              hoverinfo="text",
              text = ~paste0(
                "Phenotype: ", description,
                "<br>Round 1 Liability SNP h2: ", round(round1_h2_liability,5), " (p=",signif(round1_h2_p, 3),")",
                "<br>Round 2 Liability SNP h2: ", round(round2_h2_liability,5), " (p=",signif(round2_h2_p, 3),")",
                "<br>Round 2 Effective N: ", Neff)) %>% add_trace(
                  x=~round1_h2_liability,
                  y=~round1_h2_liability,
                  mode="lines",
                  hoverinfo="none",
                  showlegend=F
                ) %>% onRender(js)

# construct
bscols(widths=c(10,2),
       config(pp, collaborate = F, showLink=F, displayModeBar="hover", displaylogo=F, sendData=F),
       list(
      		filter_slider("Neff", "Effective N", shared_h2, ~Neff, step=100, width="100%",round=T),
      		filter_select("variable_type","Var. Type", shared_h2, ~variable_type, multiple=T),
      		filter_select("confidence","Confidence", shared_h2, ~confidence, multiple=T),
      		filter_select("source","Dataset", shared_h2, ~source, multiple=T)
       )
)
```

<p></p>

*Note:* Comparison limited to phenotypes with at least some [confidence](confidence.html) in the Round 2 results. Trend is clearer when filtering to higher effective sample size using the slider on the right.


</div>

<br>

#### SNP Heritability p-values
  
<div class="well">
  
```{r qq_h2, echo=FALSE}

shared_h2_3 <- SharedData$new(dat[dat$confidence != "none",])

exp_nlogp <- function(p){
  nn <- length(p)
  qquad <- 1:nn
  qref <- (qquad-.5)/nn
  idx <- rank(p)
  return(-log10(qref[idx]))
}

nnr <- nrow(dat[dat$confidence != "none",])
qquad <-c(1:nnr)
qref <- ((qquad-.5)/nnr)
ci_up <- qbeta(.975,qquad,nnr+1-qquad)
ci_lo <- qbeta(.025,qquad,nnr+1-qquad)

pp <- plot_ly(shared_h2_3, 
              x=~exp_nlogp(round1_h2_p), 
              y=~(-log10(round1_h2_p)),
              type="scatter",
              mode="markers",
              customdata=~urls,
              showlegend=T,
              name="Round 1",
              hoverinfo="text",
              text = ~paste0(
                "Phenotype: ", description,
                "<br>Round 1 Liability SNP h2: ", round(round1_h2_liability,5), " (p=",signif(round1_h2_p, 3),")",
                "<br>Round 2 Liability SNP h2: ", round(round2_h2_liability,5), " (p=",signif(round2_h2_p, 3),")",
                "<br>Round 2 Effective N: ", Neff)
) %>% add_trace(
  x=~exp_nlogp(round2_h2_p), 
  y=~(-log10(round2_h2_p)),
  type="scatter",
  mode="markers",
  customdata=~urls,
  showlegend=T,
  name="Round 2",
  hoverinfo="text",
  text = ~paste0(
    "Phenotype: ", description,
    "<br>Round 1 h2: ", round(round1_h2_liability,5), " (p=",signif(round1_h2_p, 3),")",
    "<br>Round 2 h2: ", round(round2_h2_liability,5), " (p=",signif(round2_h2_p, 3),")",
    "<br>Round 2Effective N: ", Neff)
) %>% add_trace(
  x=-log10(qref),
  y=-log10(qref),
  mode="lines",
  showlegend=F,
  hoverinfo = "text",
  text = ""
) %>% add_trace(
  x=-log10(qref),
  y=-log10(ci_up),
  mode="lines",
  line=list(color='#2ca02c'),
  showlegend=F,
  hoverinfo = "text",
  text = ""
) %>% add_trace(
  x=-log10(qref),
  y=-log10(ci_lo),
  mode="lines",
  line=list(color='#2ca02c'),
  fill="tonexty",
  fillcolor='rgba(44,160,44,0.2)',
  showlegend=F,
  hoverinfo = "text",
  text = ""
) %>% layout(
  showlegend=T,
  xaxis = list(title="Expected -log10(p-value)"),
  yaxis = list(title="Observed -log10(p-value)")
) %>% onRender(js)

bscols(widths=c(12),
       config(pp, collaborate = F, showLink=F, displayModeBar="hover", displaylogo=F, sendData=F)
)
```

<p></p>

*Note:* Comparison limited to phenotypes with at least some [confidence](confidence.html) in the Round 2 results.

</div>

<br>

### Precision  {.tabset .tabset-fade}

We can compare the standard errors (SEs) of the estimates between Round 1 and Round 2 to evaulate how much precision was gained from the expanded sample size in Round 2. We observe generally smaller SEs for the SNP heritability and slightly increased SEs for the intercept.

#### SNP Heritability SEs

<div class="well">

```{r comp_h2_se, echo=FALSE}
# setup shared data
shared_h2 <- SharedData$new(dat[dat$confidence != "none",])

# define plot
pp <- plot_ly(shared_h2, 
              x=~round1_h2_liability_se, 
              y=~round2_h2_liability_se, 
              type="scatter",
              mode="markers",
              customdata=~urls,
              hoverinfo="text",
              text = ~paste0(
                "Phenotype: ", description,
                "<br>Round 1 Liability SNP h2: ", round(round1_h2_liability,5), " (se=",signif(round1_h2_liability_se, 3),")",
                "<br>Round 2 Liability SNP h2: ", round(round2_h2_liability,5), " (se=",signif(round2_h2_liability_se, 3),")",
                "<br>Round 2 Effective N: ", Neff)) %>% add_trace(
                  x=~round1_h2_liability_se,
                  y=~round1_h2_liability_se,
                  mode="lines",
                  hoverinfo="none",
                  showlegend=F
                ) %>% onRender(js)

# construct
bscols(widths=c(10,2),
       config(pp, collaborate = F, showLink=F, displayModeBar="hover", displaylogo=F, sendData=F),
       list(
      		filter_slider("Neff", "Effective N", shared_h2, ~Neff, step=100, width="100%",round=T),
      		filter_select("variable_type","Var. Type", shared_h2, ~variable_type, multiple=T),
      		filter_select("confidence","Confidence", shared_h2, ~confidence, multiple=T),
      		filter_select("source","Dataset", shared_h2, ~source, multiple=T)
       )
)
```

<p></p>

*Note:* Comparison limited to phenotypes with at least some [confidence](confidence.html) in the Round 2 results.

</div>

<br>

#### Intercept SEs

<div class="well">

```{r comp_int_se, echo=FALSE}
# setup shared data
shared_int4 <- SharedData$new(dat[dat$confidence != "none",])

# define plot
pp <- plot_ly(shared_int4, 
              x=~round1_intercept_se, 
              y=~round2_intercept_se, 
              type="scatter",
              mode="markers",
              customdata=~urls,
              hoverinfo="text",
              text = ~paste0(
                "Phenotype: ", description,
                "<br>Round 1 Int.: ", round(round1_intercept,5), " (se=",signif(round1_intercept_se,3),")",
                "<br>Round 2 Int.: ", round(round2_intercept,5), " (se=",signif(round2_intercept_se,3),")",
                "<br>Round 2 Effective N: ", Neff)
            ) %>% add_trace(
                x=~round1_intercept_se,
                y=~round1_intercept_se,
                mode="lines",
                hoverinfo="none",
                showlegend=F
            ) %>% onRender(js)

# construct
bscols(widths=c(10,2),
       config(pp, collaborate = F, showLink=F, displayModeBar="hover", displaylogo=F, sendData=F),
       list(
      		filter_slider("Neff", "Effective N", shared_int4, ~Neff, step=100, width="100%",round=T),
      		filter_select("variable_type","Var. Type", shared_int4, ~variable_type, multiple=T),
      		filter_select("confidence","Confidence", shared_int4, ~confidence, multiple=T),
      		filter_select("source","Dataset", shared_int4, ~source, multiple=T)
       )
)
```

<p></p>

*Note:* Comparison limited to phenotypes with at least some [confidence](confidence.html) in the Round 2 results.

</div>

<br>
