---
title: "Visualising $h^2_g$ and intercept results"
date: "Last updated `r format(Sys.Date())`"
author: "Results from the [Neale Lab](credits.html)"
output: 
  html_document:
    toc: true
    toc_float: true
params:
  datfile: "../results/round2_final/ukb31063_h2_topline.02Oct2019.tsv.gz"
---

```{r child = '_toc_fix.Rmd'}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(reshape2)
require(ggplot2)
require(plotly)
require(DT)
require(crosstalk)
require(crosstool)
require(Rmpfr)

# handle params
dat_full <- read.delim(file=normalizePath(params$datfile),sep='\t',header=T,stringsAsFactors=F,colClasses=c(intercept_p="character"))

# setup
dat <- dat_full[dat_full$confidence != "none",]
dat$confidence <- factor(dat$confidence, levels=c("high","medium","low"))

plotly_colors <- c(
    '#1f77b4',  # muted blue
    '#ff7f0e',  # safety orange
    '#2ca02c',  # cooked asparagus green
    '#d62728',  # brick red
    '#9467bd',  # muted purple
    '#8c564b',  # chestnut brown
    '#e377c2',  # raspberry yogurt pink
    '#7f7f7f',  # middle gray
    '#bcbd22',  # curry yellow-green
    '#17becf'   # blue-teal
)

```

```{r smallp_func, include=FALSE}

# handle intercept_p as a string (due to high precision p-vals)
pstring <- function(p, prec=3){

intp_string <- as.character(p)
intp_string[p == as.character(as.numeric(p))] <- as.character(signif(as.numeric(p),prec)[(p == as.character(as.numeric(p)))])
highstr <- intp_string[!(p == as.character(as.numeric(p)))]
highstr_lead <- as.character(signif(as.numeric(substr(highstr,1,6)),prec))
highstr_exp <- sapply(highstr,function(a) strsplit(a,split="e")[[1]][2],USE.NAMES = F)
intp_string[!(p == as.character(as.numeric(p)))] <- paste0(highstr_lead,"e",highstr_exp)
while(any(intp_string=="1")){
  prec <- prec+1
  intp_string[intp_string=="1"] <- as.character(signif(as.numeric(p),prec)[intp_string=="1"])
}
return(intp_string)
}


dat$intercept_p <- pstring(dat$intercept_p)
dat$int_nlogp <- as.numeric(-log10(mpfr(0.5,64)*erfc(mpfr(dat$intercept_z,64)/sqrt(mpfr(2,64)))))

```

<br>

### Overview

<div class="well">

The plots here are designed to allow browsing overall trends in the SNP heritability and intercepts results from LD Score regression of the [Neale Lab GWAS](https://www.nealelab.is/uk-biobank). Individual phenotype results can be viewed in the [browser](h2_browser.html). 

All plots below are restricted to results with at least some confidence in the LDSR results, and report SNP heritability on the liabulit scale. See the [Methods](details.html) section for more information on our [selection of GWAS phenotypes](select_topline.html), [confidence ratings](confidence.html) for the LDSR results, and [significance thresholds](significance.html).

In addition to the plots below, additional visualizations are available for considering:

* [Partitioned SNP heritability results](viz_annot.html)
* [The impact of sample size](viz_sampsize.html)
* [Comparison of these results to the Round 1 GWAS](viz_R1vR2.html)

Previous plots from the Round 1 results can be found [in the Archive section](round1_plots_home.html).

</div>

<br>

### SNP Heritability {.tabset .tabset-fade}

The average SNP hertiability estimate across all `r nrow(dat_full)` primary GWAS is `r round(mean(dat_full$h2_liability),3)` (median = `r round(median(dat_full$h2_liability),3)`). Average estimates increase when restricting to any [confidence](confidence.html) (mean = `r round(mean(dat$h2_liability),3)`, median = `r round(median(dat$h2_liability),3)`) or `high` confidence (mean = `r round(mean(dat$h2_liability[dat$confidence=="high"]),3)`, median = `r round(median(dat$h2_liability[dat$confidence=="high"]),3)`). Strong departure from the null hypothesis of $h^2_g=0$ is observed across phenotypes, especially among high confidence results.

<p></p>

#### SNP Heritability Estimates

<div class="well">

```{r hist_h2, echo=FALSE}
d2 <- SharedData$new(dat)

pp <- plot_ly(d2) %>%
        add_histogram(
			    x = ~h2_liability,
			    histnorm = "probability",
			    hoverinfo="none"
	  		)
  
  
bscols(widths=c(10,2),
	config(pp, collaborate = F, showLink=F, displayModeBar=F, displaylogo=F, sendData=F),
	list(
		filter_slider("Neff", "Effective N", d2, ~Neff, step=100, width="100%"),
		filter_select("variable_type","Var. Type", d2, ~variable_type, multiple=T),
		filter_select("confidence","Confidence", d2, ~confidence, multiple=T),
		filter_select("significance","h2 Significance", d2, ~h2_sig, multiple=T),
		filter_select("source","Dataset", d2, ~source, multiple=T)
	)
)
```

</div>

<br>

#### SNP Heritability p-values

<div class="well">
  
```{r qq_h2, echo=FALSE}

shared_h2_3 <- SharedData$new(dat)

exp_nlogp <- function(p){
  nn <- length(p)
  qquad <- 1:nn
  qref <- (qquad-.5)/nn
  idx <- rank(p)
  return(-log10(qref[idx]))
}

nnr <- nrow(dat[dat$confidence != "none",])
qquad <-c(1:nnr)
qref <- ((qquad-.5)/nnr)
ci_up <- qbeta(.975,qquad,nnr+1-qquad)
ci_lo <- qbeta(.025,qquad,nnr+1-qquad)

pp <- plot_ly(shared_h2_3, 
              x=~exp_nlogp(h2_p), 
              y=~(-log10(h2_p)),
              type="scatter",
              mode="markers",
              showlegend=F,
              hoverinfo="text",
              text = ~paste0(
                "Phenotype: ", description,
                "<br>SNP h2: ", round(h2_liability,5), " (p=",signif(h2_p, 3),")",
                "<br>Intercept: ", round(intercept,5), " (p=",intercept_p,")",                
                "<br>Effective N: ", Neff)
) %>% add_trace(
		x=-log10(qref),
		y=-log10(qref),
		type="scatter",
		mode="lines",
		line=list(color='rgba(0,0,0,0.8)'),
		showlegend=F,
		hoverinfo = "text",
		text = ""
) %>% add_trace(
		x=-log10(qref),
		y=-log10(ci_up),
		type="scatter",
		mode="lines",
		line=list(color='rgba(0,0,0,0.3)'),
		showlegend=F,
		hoverinfo = "text",
		text = ""
) %>% add_trace(
		x=-log10(qref),
		y=-log10(ci_lo),
		type="scatter",
		mode="lines",
		line=list(color='rgba(0,0,0,0.3)'),
		fill="tonexty",
		fillcolor='rgba(0,0,0,0.2)',
		showlegend=F,
		hoverinfo = "text",
		text = ""
) %>% layout(
  showlegend=T,
  xaxis = list(title="Expected -log10(p-value)"),
  yaxis = list(title="Observed -log10(p-value)")
)

bscols(widths=c(12),
       config(pp, collaborate = F, showLink=F, displayModeBar=F, displaylogo=F, sendData=F)
)
```

<p></p>

*Note:* The displayed 95% confidence bands assume tests are independent across phenotypes and do not account for multiple testing across annotations.

</div>

<br>

#### p-values by Confidence

<div class="well">
  
```{r qq_h2_conf, echo=FALSE}

exp_nlogp <- function(p){
  nn <- length(p)
  qquad <- 1:nn
  qref <- (qquad-.5)/nn
  idx <- rank(p)
  return(-log10(qref[idx]))
}

nnr <- nrow(dat[dat$confidence != "none",])
qquad <-c(1:nnr)
qref <- ((qquad-.5)/nnr)
ci_up <- qbeta(.975,qquad,nnr+1-qquad)
ci_lo <- qbeta(.025,qquad,nnr+1-qquad)

pp <- plot_ly(dat) %>%
        add_trace(
              data=dat[dat$confidence=="high",],
              x=~exp_nlogp(h2_p), 
              y=~(-log10(h2_p)),
              name="High Confidence",
              type="scatter",
              mode="markers",
              hoverinfo="text",
              text = ~paste0(
                "Phenotype: ", description,
                "<br>SNP h2: ", round(h2_liability,5), " (p=",signif(h2_p, 3),")",
                "<br>Intercept: ", round(intercept,5), " (p=",intercept_p,")",                
                "<br>Effective N: ", Neff)
) %>% add_trace(
              data=dat[dat$confidence=="medium",],
              x=~exp_nlogp(h2_p), 
              y=~(-log10(h2_p)),
              name="Medium Confidence",
              type="scatter",
              mode="markers",
              hoverinfo="text",
              text = ~paste0(
                "Phenotype: ", description,
                "<br>SNP h2: ", round(h2_liability,5), " (p=",signif(h2_p, 3),")",
                "<br>Intercept: ", round(intercept,5), " (p=",intercept_p,")",                
                "<br>Effective N: ", Neff)
) %>% add_trace(
              data=dat[dat$confidence=="low",],
              x=~exp_nlogp(h2_p), 
              y=~(-log10(h2_p)),
              name="Low Confidence",
              type="scatter",
              mode="markers",
              hoverinfo="text",
              text = ~paste0(
                "Phenotype: ", description,
                "<br>SNP h2: ", round(h2_liability,5), " (p=",signif(h2_p, 3),")",
                "<br>Intercept: ", round(intercept,5), " (p=",intercept_p,")",                
                "<br>Effective N: ", Neff)
) %>% add_trace(
		x=-log10(qref),
		y=-log10(qref),
		type="scatter",
		mode="lines",
		line=list(color='rgba(0,0,0,0.8)'),
		showlegend=F,
		hoverinfo = "text",
		text = ""
) %>% add_trace(
		x=-log10(qref),
		y=-log10(ci_up),
		type="scatter",
		mode="lines",
		line=list(color='rgba(0,0,0,0.3)'),
		showlegend=F,
		hoverinfo = "text",
		text = ""
) %>% add_trace(
		x=-log10(qref),
		y=-log10(ci_lo),
		type="scatter",
		mode="lines",
		line=list(color='rgba(0,0,0,0.3)'),
		fill="tonexty",
		fillcolor='rgba(0,0,0,0.2)',
		showlegend=F,
		hoverinfo = "text",
		text = ""
) %>% layout(
  showlegend=T,
  xaxis = list(title="Expected -log10(p-value)"),
  yaxis = list(title="Observed -log10(p-value)"),
  legend = list(traceorder="normal")
)

bscols(widths=c(12),
       config(pp, collaborate = F, showLink=F, displayModeBar=F, displaylogo=F, sendData=F)
)
```

<p></p>

*Note:* The displayed 95% confidence bands assume tests are independent across phenotypes within confidence rating and do not account for multiple testing across annotations. Read more about the confidence ratings [here](confidence.html).

</div>

<br>

### LDSR Intercept {.tabset .tabset-fade}

Relative GWAS confounding or LDSR model misspecification is limited across the phenotypes, with a median intercept ratio of `r median(dat_full$ratio)`. Intercepts remain statistically significant for many phenotypes, including many `high` confidence phenotypes. Outlier intercept values > 1.2 are observed for home and birth location, height, and certain biomarkers.

<p></p>

#### Ratio Estimates

<div class="well">

```{r hist_rat, echo=FALSE}
d2 <- SharedData$new(dat[dat$mean_chi2 > 1,])

pp <- plot_ly(d2) %>%
        add_histogram(
			    x = ~ratio,
			    histnorm = "probability",
			    hoverinfo="none"
	  		)
  
  
bscols(widths=c(10,2),
	config(pp, collaborate = F, showLink=F, displayModeBar=F, displaylogo=F, sendData=F),
	list(
		filter_slider("Neff", "Effective N", d2, ~Neff, step=100, width="100%"),
		filter_select("variable_type","Var. Type", d2, ~variable_type, multiple=T),
		filter_select("confidence","Confidence", d2, ~confidence, multiple=T),
		filter_select("source","Dataset", d2, ~source, multiple=T)
	)
)
```

<p></p>

*Note:* Ratio estimates restricted to phenotypes with mean $\chi^2 > 1$. Clearer results are observed by restricting to `high` confidence results and/or higher effective sample size using the filter on the right. The LDSR ratio estimate is:

$$\frac{\text{intercept}-1}{\text{mean}(\chi^2)-1}$$

</div>

<br>

#### Intercept Estimates

<div class="well">

```{r hist_int, echo=FALSE}
d2 <- SharedData$new(dat)

pp <- plot_ly(d2) %>%
        add_histogram(
			    x = ~intercept,
			    histnorm = "probability",
			    hoverinfo="none"
	  		)
  
  
bscols(widths=c(10,2),
	config(pp, collaborate = F, showLink=F, displayModeBar=F, displaylogo=F, sendData=F),
	list(
		filter_slider("Neff", "Effective N", d2, ~Neff, step=100, width="100%"),
		filter_select("variable_type","Var. Type", d2, ~variable_type, multiple=T),
		filter_select("confidence","Confidence", d2, ~confidence, multiple=T),
		filter_select("source","Dataset", d2, ~source, multiple=T)
	)
)
```

</div>

<br>

#### Int. p-values

<div class="well">
  
```{r qq_int, echo=FALSE}

shared_h2_3 <- SharedData$new(dat)

exp_nlogp <- function(p){
  nn <- length(p)
  qquad <- 1:nn
  qref <- (qquad-.5)/nn
  idx <- rank(p)
  return(-log10(qref[idx]))
}

nnr <- nrow(dat[dat$confidence != "none",])
qquad <-c(1:nnr)
qref <- ((qquad-.5)/nnr)
ci_up <- qbeta(.975,qquad,nnr+1-qquad)
ci_lo <- qbeta(.025,qquad,nnr+1-qquad)

pp <- plot_ly(shared_h2_3, 
              x=~exp_nlogp(-intercept_z), 
              y=~int_nlogp,
              type="scatter",
              mode="markers",
              showlegend=F,
              hoverinfo="text",
              text = ~paste0(
                "Phenotype: ", description,
                "<br>SNP h2: ", round(h2_liability,5), " (p=",signif(h2_p, 3),")",
                "<br>Intercept: ", round(intercept,5), " (p=",intercept_p,")",                
                "<br>Effective N: ", Neff)
) %>% add_trace(
		x=-log10(qref),
		y=-log10(qref),
		type="scatter",
		mode="lines",
		line=list(color='rgba(0,0,0,0.8)'),
		showlegend=F,
		hoverinfo = "text",
		text = ""
) %>% add_trace(
		x=-log10(qref),
		y=-log10(ci_up),
		type="scatter",
		mode="lines",
		line=list(color='rgba(0,0,0,0.3)'),
		showlegend=F,
		hoverinfo = "text",
		text = ""
) %>% add_trace(
		x=-log10(qref),
		y=-log10(ci_lo),
		type="scatter",
		mode="lines",
		line=list(color='rgba(0,0,0,0.3)'),
		fill="tonexty",
		fillcolor='rgba(0,0,0,0.2)',
		showlegend=F,
		hoverinfo = "text",
		text = ""
) %>% layout(
  showlegend=T,
  xaxis = list(title="Expected -log10(p-value)"),
  yaxis = list(title="Observed -log10(p-value)")
)

bscols(widths=c(12),
       config(pp, collaborate = F, showLink=F, displayModeBar=F, displaylogo=F, sendData=F)
)
```


<p></p>

*Note:* The displayed 95% confidence bands assume tests are independent across phenotypes and do not account for multiple testing across annotations.

</div>

<br>

#### Int. p-values by Confidence

<div class="well">
  
```{r qq_int_conf, echo=FALSE}

exp_nlogp <- function(p){
  nn <- length(p)
  qquad <- 1:nn
  qref <- (qquad-.5)/nn
  idx <- rank(p)
  return(-log10(qref[idx]))
}

nnr <- nrow(dat[dat$confidence != "none",])
qquad <-c(1:nnr)
qref <- ((qquad-.5)/nnr)
ci_up <- qbeta(.975,qquad,nnr+1-qquad)
ci_lo <- qbeta(.025,qquad,nnr+1-qquad)

pp <- plot_ly(dat) %>%
        add_trace(
              data=dat[dat$confidence=="high",],
              x=~exp_nlogp(-intercept_z), 
              y=~int_nlogp,
              name="High Confidence",
              type="scatter",
              mode="markers",
              hoverinfo="text",
              text = ~paste0(
                "Phenotype: ", description,
                "<br>SNP h2: ", round(h2_liability,5), " (p=",signif(h2_p, 3),")",
                "<br>Intercept: ", round(intercept,5), " (p=",intercept_p,")",                
                "<br>Effective N: ", Neff)
) %>% add_trace(
              data=dat[dat$confidence=="medium",],
              x=~exp_nlogp(-intercept_z), 
              y=~int_nlogp,
              name="Medium Confidence",
              type="scatter",
              mode="markers",
              hoverinfo="text",
              text = ~paste0(
                "Phenotype: ", description,
                "<br>SNP h2: ", round(h2_liability,5), " (p=",signif(h2_p, 3),")",
                "<br>Intercept: ", round(intercept,5), " (p=",intercept_p,")",                
                "<br>Effective N: ", Neff)
) %>% add_trace(
              data=dat[dat$confidence=="low",],
              x=~exp_nlogp(-intercept_z), 
              y=~int_nlogp,
              name="Low Confidence",
              type="scatter",
              mode="markers",
              hoverinfo="text",
              text = ~paste0(
                "Phenotype: ", description,
                "<br>SNP h2: ", round(h2_liability,5), " (p=",signif(h2_p, 3),")",
                "<br>Intercept: ", round(intercept,5), " (p=",intercept_p,")",                
                "<br>Effective N: ", Neff)
) %>% add_trace(
		x=-log10(qref),
		y=-log10(qref),
		type="scatter",
		mode="lines",
		line=list(color='rgba(0,0,0,0.8)'),
		showlegend=F,
		hoverinfo = "text",
		text = ""
) %>% add_trace(
		x=-log10(qref),
		y=-log10(ci_up),
		type="scatter",
		mode="lines",
		line=list(color='rgba(0,0,0,0.3)'),
		showlegend=F,
		hoverinfo = "text",
		text = ""
) %>% add_trace(
		x=-log10(qref),
		y=-log10(ci_lo),
		type="scatter",
		mode="lines",
		line=list(color='rgba(0,0,0,0.3)'),
		fill="tonexty",
		fillcolor='rgba(0,0,0,0.2)',
		showlegend=F,
		hoverinfo = "text",
		text = ""
) %>% layout(
  showlegend=T,
  xaxis = list(title="Expected -log10(p-value)"),
  yaxis = list(title="Observed -log10(p-value)")
)

bscols(widths=c(12),
       config(pp, collaborate = F, showLink=F, displayModeBar=F, displaylogo=F, sendData=F)
)
```

<p></p>

*Note:* The displayed 95% confidence bands assume tests are independent across phenotypes within confidence rating and do not account for multiple testing across annotations. Read more about the confidence ratings [here](confidence.html).

</div>

<br>

### SNP Heritablity vs. Intercept {.tabset .tabset-fade}

Among `medium` and `high` confidence phenotypes, there's a positive relationship between the $h^2_g$ estimate and the intercept estimate. This trend may indicate some degree of model misspecification in the paritioned LDSR model. Sample size may also be a factor if $h^2_g$ is biased at some sample sizes, but comparison of $h^2_g$ to the intercept ratio shows a similar trend towards non-zero ratios where $h^2_g$ is large.

#### Estimates

<div class="well">

```{r h2_int, echo=FALSE}
d2 <- SharedData$new(dat)

pp <- plot_ly(d2, 
              x=~intercept, 
              y=~h2_liability,
              type="scatter",
              mode="markers",
              color=~confidence,
              colors=plotly_colors[1:3],
              showlegend=T,
              hoverinfo="text",
              text = ~paste0(
                "Phenotype: ", description,
                "<br>SNP h2: ", round(h2_liability,5), " (p=",signif(h2_p, 3),")",
                "<br>Intercept: ", round(intercept,5), " (p=",intercept_p,")",                
                "<br>Effective N: ", Neff)
)

bscols(widths=c(10,2),
	config(pp, collaborate = F, showLink=F, displayModeBar=F, displaylogo=F, sendData=F),
	list(
		filter_slider("Neff", "Effective N", d2, ~Neff, step=100, width="100%"),
		filter_select("variable_type","Var. Type", d2, ~variable_type, multiple=T),
		filter_select("confidence","Confidence", d2, ~confidence, multiple=T),
		filter_select("source","Dataset", d2, ~source, multiple=T)
	)
)
```

<p></p>

*Note:* Clearer results are observed by restricting to `high` confidence results, continuous variables, and/or higher effective sample size using the filter on the right. 

</div>

<br>

#### Z scores

<div class="well">

```{r h2_int_z, echo=FALSE}
d3 <- SharedData$new(dat)

pp <- plot_ly(d3, 
              x=~intercept_z, 
              y=~h2_z,
              type="scatter",
              mode="markers",
              color=~confidence,
              colors=plotly_colors[1:3],
              showlegend=T,
              hoverinfo="text",
              text = ~paste0(
                "Phenotype: ", description,
                "<br>SNP h2: ", round(h2_liability,5), " (p=",signif(h2_p, 3),")",
                "<br>Intercept: ", round(intercept,5), " (p=",intercept_p,")",                
                "<br>Effective N: ", Neff)
)

bscols(widths=c(10,2),
	config(pp, collaborate = F, showLink=F, displayModeBar=F, displaylogo=F, sendData=F),
	list(
		filter_slider("Neff", "Effective N", d3, ~Neff, step=100, width="100%"),
		filter_select("variable_type","Var. Type", d3, ~variable_type, multiple=T),
		filter_select("confidence","Confidence", d3, ~confidence, multiple=T),
		filter_select("source","Dataset", d3, ~source, multiple=T)
	)
)
```

</div>

<br>

#### SNP Heritability vs. Ratio

<div class="well">

```{r h2_ratio, echo=FALSE}
d4 <- SharedData$new(dat[dat$mean_chi2 > 1,])

pp <- plot_ly(d4, 
              x=~ratio, 
              y=~h2_liability,
              type="scatter",
              mode="markers",
              color=~confidence,
              colors=plotly_colors[1:3],
              showlegend=T,
              hoverinfo="text",
              text = ~paste0(
                "Phenotype: ", description,
                "<br>SNP h2: ", round(h2_liability,5), " (p=",signif(h2_p, 3),")",
                "<br>Intercept: ", round(intercept,5), " (p=",intercept_p,")",                
                "<br>Effective N: ", Neff)
)

bscols(widths=c(10,2),
	config(pp, collaborate = F, showLink=F, displayModeBar=F, displaylogo=F, sendData=F),
	list(
		filter_slider("Neff", "Effective N", d4, ~Neff, step=100, width="100%"),
		filter_select("variable_type","Var. Type", d4, ~variable_type, multiple=T),
		filter_select("confidence","Confidence", d4, ~confidence, multiple=T),
		filter_select("source","Dataset", d4, ~source, multiple=T)
	)
)
```

<p></p>

*Note:* Ratio estimates restricted to phenotypes with mean $\chi^2 > 1$. Clearer results are observed by restricting to `high` confidence results, continuous phenotypes, and/or higher effective sample sizes using the filter on the right. The LDSR ratio estimate is:

$$\frac{\text{intercept}-1}{\text{mean}(\chi^2)-1}$$

</div>

<br>

