---
title: "Visualizing $h^2$ results: Univariate LDSR"
date: "Last updated `r format(Sys.Date())`"
output: 
  html_document:
    toc: true
    toc_float: true
params:
  datfile1: "../results/ukbb_h2part.RData"
  datfile2: "../results/ukbb_h2univar.RData"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(reshape2)
require(ggplot2)
require(plotly)
require(DT)
require(crosstalk)
require(crosstool)

# handle params
load(file=params$datfile1)
dat1 <- dat[,c("phenotype","description","N","N_case","N_control","prevelence","mean_chi2","lambdaGC","intercept","intercept_z","intercept_p","ratio","h2_liability","h2_z","h2_p")]
names(dat1) <- gsub("intercept","partitioned_intercept",names(dat1))
names(dat1) <- gsub("h2","partitioned_h2",names(dat1))
names(dat1) <- gsub("ratio","partitioned_ratio",names(dat1))

load(file=params$datfile2)
dat2 <- dat[,c("phenotype","intercept","intercept_z","intercept_p","ratio","h2_liability","h2_z","h2_p")]
names(dat2) <- gsub("intercept","univariate_intercept",names(dat2))
names(dat2) <- gsub("h2","univariate_h2",names(dat2))
names(dat2) <- gsub("ratio","univariate_ratio",names(dat2))

rm(dat)
dat <- merge(dat1,dat2,by="phenotype")

# setup
dat$isBinary <- !is.na(dat$N_case)
dat$Neff <- dat$N
dat$Neff[dat$isBinary] <- round( (4/((1/dat$N_case)+(1/dat$N_control)))[dat$isBinary], 2)

dat <- dat[dat$Neff > 200,]

dat$table_desc <- paste0("<a href='h2_summary_",dat$phenotype,".html'>",dat$description,"</a>")

```

<br>

### Overview

<div class="well">

Comparison of our primary $h^2$ results from partitioned LDSR to results from univariate LDSR with default settings. These differ in two key ways:

* **Reference panel:** Univariate 1000 Genomes European LD scores (i.e. precomputed `./eur_w_ld`) vs. BaselineLD v1.1 LD scores from 1000 Genomes Europeans, as described in [Gazal et al. 2017](http://www.nature.com/ng/journal/vaop/ncurrent/full/ng.3954.html) (also on [biorxiv](http://www.biorxiv.org/content/early/2017/04/03/082024))
* **Max chi^2 filtering:** Default settings with "two-step" estimation (limiting intercept estimation to regression with $\chi^2 < 30$ before estimating $h^2$ in regression with the intercept fixed to the first stage estimate) vs. one-step estimation with no $\chi^2$ filter

</div>

<br>

### Intercept Comparison {.tabset .tabset-fade}

#### Estimates

<div class="well">

```{r comp_int, echo=FALSE}
# setup shared data
shared_int <- SharedData$new(dat)

# define plot
pp <- plot_ly(shared_int, 
			  x=~univariate_intercept, 
			  y=~partitioned_intercept, 
			  type="scatter",
			  mode="markers",
			  hoverinfo="text",
			  text = ~paste0(
			  		    "Phenotype: ", description,
			  			"<br>Univariate Int.: ", round(univariate_intercept,5), " (p=",signif(univariate_intercept_p, 3),")",
			  			"<br>Partitioned Int.: ", round(partitioned_intercept,5), " (p=",signif(partitioned_intercept_p, 3),")",
			  			"<br>Effective N: ", Neff)) %>% add_trace(
			  			x=~univariate_intercept,
			  			y=~univariate_intercept,
			  			mode="lines",
			  			hoverinfo="none",
			  			showlegend=F
			  			)

# construct
bscols(widths=c(10,2),
	config(pp, collaborate = F, showLink=F, displayModeBar=F, displaylogo=F, sendData=F),
	list(
		filter_slider("Neff", "Effective N", shared_int, ~Neff, step=100, width="100%"),
		filter_select("isBinary","Binary Pheno.", shared_int, ~isBinary, multiple=T)
	)
)
```
</div>

#### Ratio of Estimates

<div class="well">

```{r hist_int, echo=FALSE}
shared_int2 <- SharedData$new(dat)

pp <- plot_ly(shared_int2,
			  x = ~(partitioned_intercept/univariate_intercept), 
			  type = "histogram",
			  histnorm = "probability"
	  		) %>% layout(xaxis=list(title="Partitioned Int. / Univariate Int."), yaxis=list(title="Frequency"))

bscols(widths=c(10,2),
	config(pp, collaborate = F, showLink=F, displayModeBar=F, displaylogo=F, sendData=F),
	list(
		filter_slider("Neff", "Effective N", shared_int2, ~Neff, step=100, width="100%"),
		filter_select("isBinary","Binary Pheno.", shared_int2, ~isBinary, multiple=T)
	)
)
```

</div>


#### Significance

<div class="well">

```{r qq_int, echo=FALSE}

shared_int3 <- SharedData$new(dat)

exp_nlogp <- function(p){
	nn <- length(p)
	qquad <- 1:nn
	qref <- (qquad-.5)/nn
	idx <- rank(p)
	return(-log10(qref[idx]))
}

qquad <-c(1:nrow(dat))
qref <- ((qquad-.5)/nrow(dat))
ci_up <- qbeta(.975,qquad,nrow(dat)+1-qquad)
ci_lo <- qbeta(.025,qquad,nrow(dat)+1-qquad)

pp <- plot_ly(shared_int3, 
			  x=~exp_nlogp(univariate_intercept_p), 
			  y=~(-log10(univariate_intercept_p)),
			  type="scatter",
			  mode="markers",
			  showlegend=T,
			  name="Univariate",
			  hoverinfo="text",
			  text = ~paste0(
			  		    "Phenotype: ", description,
			  			"<br>Univariate Int.: ", round(univariate_intercept,5), " (p=",signif(univariate_intercept_p, 3),")",
			  			"<br>Partitioned Int.: ", round(partitioned_intercept,5), " (p=",signif(partitioned_intercept_p, 3),")",
			  			"<br>Effective N: ", Neff)
	) %>% add_trace(
		x=~exp_nlogp(partitioned_intercept_p), 
			  y=~(-log10(partitioned_intercept_p)),
			  type="scatter",
			  mode="markers",
			  showlegend=T,
			  name="Partitioned",
			  hoverinfo="text",
			  text = ~paste0(
			  		    "Phenotype: ", description,
			  			"<br>Univariate Int.: ", round(univariate_intercept,5), " (p=",signif(univariate_intercept_p, 3),")",
			  			"<br>Partitioned Int.: ", round(partitioned_intercept,5), " (p=",signif(partitioned_intercept_p, 3),")",
			  			"<br>Effective N: ", Neff)
	) %>% add_trace(
		x=-log10(qref),
		y=-log10(qref),
		mode="lines",
		showlegend=F,
		hoverinfo = "text",
		text = ""
	) %>% add_trace(
		x=-log10(qref),
		y=-log10(ci_up),
		mode="lines",
		line=list(color='#2ca02c'),
		showlegend=F,
		hoverinfo = "text",
		text = ""
	) %>% add_trace(
		x=-log10(qref),
		y=-log10(ci_lo),
		mode="lines",
		line=list(color='#2ca02c'),
		fill="tonexty",
		fillcolor='rgba(44,160,44,0.2)',
		showlegend=F,
		hoverinfo = "text",
		text = ""
	) %>% layout(
	  			 showlegend=T,
	  			 xaxis = list(title="Expected -log10(p-value)"),
	  			 yaxis = list(title="Observed -log10(p-value)")
	  )

bscols(widths=c(12),
	config(pp, collaborate = F, showLink=F, displayModeBar=F, displaylogo=F, sendData=F)
)
```

</div>

<br>

### $h^2$ Comparison {.tabset .tabset-fade}

#### Estimates

<div class="well">

```{r comp_h2, echo=FALSE}
# setup shared data
shared_h2 <- SharedData$new(dat)

# define plot
pp <- plot_ly(shared_h2, 
			  x=~univariate_h2_liability, 
			  y=~partitioned_h2_liability, 
			  type="scatter",
			  mode="markers",
			  hoverinfo="text",
			  text = ~paste0(
			  		    "Phenotype: ", description,
			  			"<br>Univariate h2: ", round(univariate_h2_liability,5), " (p=",signif(univariate_h2_p, 3),")",
			  			"<br>Partitioned h2: ", round(partitioned_h2_liability,5), " (p=",signif(partitioned_h2_p, 3),")",
			  			"<br>Effective N: ", Neff)) %>% add_trace(
			  			x=~univariate_h2_liability,
			  			y=~univariate_h2_liability,
			  			mode="lines",
			  			hoverinfo="none",
			  			showlegend=F
			  			)

# construct
bscols(widths=c(10,2),
	config(pp, collaborate = F, showLink=F, displayModeBar=F, displaylogo=F, sendData=F),
	list(
		filter_slider("Neff", "Effective N", shared_h2, ~Neff, step=100, width="100%"),
		filter_select("isBinary","Binary Pheno.", shared_h2, ~isBinary, multiple=T)
	)
)
```

<br>

**Note:** Restrict to effective N > 10,000 for a clearer pattern of results.


</div>

#### Ratio of Estimates

<div class="well">

```{r hist_h2, echo=FALSE}
shared_h2_2 <- SharedData$new(dat[dat$Neff > 10000,])

pp <- plot_ly(shared_h2_2,
			  x = ~(partitioned_h2_liability/univariate_h2_liability), 
			  type = "histogram",
			  histnorm = "probability"
	  		) %>% layout(xaxis=list(title="Partitioned h2 / Univariate h2"), yaxis=list(title="Frequency"))

bscols(widths=c(10,2),
	config(pp, collaborate = F, showLink=F, displayModeBar=F, displaylogo=F, sendData=F),
	list(
		filter_slider("Neff", "Effective N", shared_h2_2, ~Neff, step=100, width="100%"),
		filter_select("isBinary","Binary Pheno.", shared_h2_2, ~isBinary, multiple=T)
	)
)
```

<br>

**Note:** To improve resolution this plot is restricted to phenotypes with effective N > 10,000.

</div>

#### Significance

<div class="well">

```{r qq_h2, echo=FALSE}

shared_h2_3 <- SharedData$new(dat)

exp_nlogp <- function(p){
	nn <- length(p)
	qquad <- 1:nn
	qref <- (qquad-.5)/nn
	idx <- rank(p)
	return(-log10(qref[idx]))
}

qquad <-c(1:nrow(dat))
qref <- ((qquad-.5)/nrow(dat))
ci_up <- qbeta(.975,qquad,nrow(dat)+1-qquad)
ci_lo <- qbeta(.025,qquad,nrow(dat)+1-qquad)

pp <- plot_ly(shared_h2_3, 
			  x=~exp_nlogp(univariate_h2_p), 
			  y=~(-log10(univariate_h2_p)),
			  type="scatter",
			  mode="markers",
			  showlegend=T,
			  name="Univariate",
			  hoverinfo="text",
			  text = ~paste0(
			  		    "Phenotype: ", description,
			  			"<br>Univariate h2: ", round(univariate_h2_liability,5), " (p=",signif(univariate_h2_p, 3),")",
			  			"<br>Partitioned h2: ", round(partitioned_h2_liability,5), " (p=",signif(partitioned_h2_p, 3),")",
			  			"<br>Effective N: ", Neff)
	) %>% add_trace(
		x=~exp_nlogp(partitioned_h2_p), 
			  y=~(-log10(partitioned_h2_p)),
			  type="scatter",
			  mode="markers",
			  showlegend=T,
			  name="Partitioned",
			  hoverinfo="text",
			  text = ~paste0(
			  		    "Phenotype: ", description,
			  			"<br>Univariate h2: ", round(univariate_h2_liability,5), " (p=",signif(univariate_h2_p, 3),")",
			  			"<br>Partitioned h2: ", round(partitioned_h2_liability,5), " (p=",signif(partitioned_h2_p, 3),")",
			  			"<br>Effective N: ", Neff)
	) %>% add_trace(
		x=-log10(qref),
		y=-log10(qref),
		mode="lines",
		showlegend=F,
		hoverinfo = "text",
		text = ""
	) %>% add_trace(
		x=-log10(qref),
		y=-log10(ci_up),
		mode="lines",
		line=list(color='#2ca02c'),
		showlegend=F,
		hoverinfo = "text",
		text = ""
	) %>% add_trace(
		x=-log10(qref),
		y=-log10(ci_lo),
		mode="lines",
		line=list(color='#2ca02c'),
		fill="tonexty",
		fillcolor='rgba(44,160,44,0.2)',
		showlegend=F,
		hoverinfo = "text",
		text = ""
	) %>% layout(
	  			 showlegend=T,
	  			 xaxis = list(title="Expected -log10(p-value)"),
	  			 yaxis = list(title="Observed -log10(p-value)")
	  )

bscols(widths=c(12),
	config(pp, collaborate = F, showLink=F, displayModeBar=F, displaylogo=F, sendData=F)
)
```

</div>

<br>
