---
title: "Vizualizing $h^2$ results: Partitioned Annotations"
date: "Last updated `r format(Sys.Date())`"
output: 
  html_document:
    toc: true
    toc_float: true
params:
  datfile: "../results/ukbb_h2part.RData"
---

```{r child = '_toc_fix.Rmd'}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(reshape2)
require(ggplot2)
require(plotly)
require(DT)
require(crosstalk)
require(crosstool)

# handle params
load(file=params$datfile)

# setup
dat$isBinary <- !is.na(dat$N_case)
dat$Neff <- dat$N
dat$Neff[dat$isBinary] <- round( (4/((1/dat$N_case)+(1/dat$N_control)))[dat$isBinary], 2)

dat <- dat[dat$Neff > 200,]

dat$isNomSig_h2 <- dat$h2_p < .05
dat$isBonfSig_h2 <- dat$h2_p < (.05/nrow(dat))
dat$isNomSig_int <- dat$intercept_p < .05
dat$isBonfSig_int <- dat$intercept_p < (.05/nrow(dat))
dat$table_desc <- paste0("<a href='h2_summary_",dat$phenotype,".html'>",dat$description,"</a>")

```

<br>

### Binary Annotations {.tabset .tabset-fade}

#### Coefficients

<div class="well">

```{r hist_coef, echo=FALSE}

dtmp <- dat
dtmp <- dtmp[,-grep("_se",names(dtmp))]
dtmp <- dtmp[,-grep("_z",names(dtmp))]
dtmp <- dtmp[,-grep("_p",names(dtmp))]
dtmp <- dtmp[,-grep("Prop_SNPs",names(dtmp))]
dtmp <- dtmp[,-grep("Prop_h2",names(dtmp))]
dtmp <- dtmp[,-grep("Enrichment",names(dtmp))]
dtmp <- dtmp[,-grep("base",names(dtmp))]
dtmp <- dtmp[,-grep("MAFbin",names(dtmp))]
dtmp <- dtmp[,-grep("CpG_Content",names(dtmp))]
dtmp <- dtmp[,-grep("GERP.NS",names(dtmp))]
dtmp <- dtmp[,-grep("MAF_Adj",names(dtmp))]
dtmp <- dtmp[,-grep("Nucleotide",names(dtmp))]
dtmp <- dtmp[,-grep("Backgrd",names(dtmp))]
dtmp <- dtmp[,-grep("Recomb_Rate",names(dtmp))]
dtmp <- dtmp[,-grep("extend.500",names(dtmp))]
dtmp <- dtmp[,!(names(dtmp) %in% c("description","N","N_case","N_control","prevelence","mean_chi2","lambdaGC","intercept","ratio","h2_observed","isNomSig_h2","isNomSig_int","table_desc","h2_liability"))]

dd <- melt(dtmp,c("phenotype","Neff","isBinary","isBonfSig_h2","isBonfSig_int"),variable.name = "annot1",value.name="Coefficient")
dd$annot <- gsub("..Coefficient","",dd$annot1)
d3 <- SharedData$new(dd[dd$isBonfSig_h2 & dd$Neff > 100000,])

pp <- plot_ly(d3, height=600) %>%
		add_trace(
			  y = ~Coefficient,
			  split = ~annot,
			  type="box",
			  hoverinfo = "x"
	  		) %>%
		layout(showlegend=F, yaxis=list(range=c(-1.3e-6,1.1e-6),exponentformat="e"), margin=list(b=150,l=100), xaxis=list(title="Annotation", tickangle=-30))

bscols(widths=c(10,2),
	config(pp, collaborate = F, showLink=F, displayModeBar=F, displaylogo=F, sendData=F),
	list(
		filter_slider("Neff", "Effective N", d3, ~Neff, step=100, width="100%"),
		filter_select("isBinary","Binary Pheno.", d3, ~isBinary, multiple=T),
		filter_select("isBonfSig_int","Signif. Int.", d3, ~isBonfSig_int, multiple=T)
	)
)
```

<br>

**Note:** For clarity, this plot is limited to phenotypes with significant $h^2$ (after Bonferroni correction) and effective N > 100,000.

</div>

<br>


#### Coefficient Z-score

<div class="well">

```{r hist_coef_z, echo=FALSE}

dtmp3 <- dat
dtmp3 <- dtmp3[,-grep("_se",names(dtmp3))]
dtmp3 <- dtmp3[,-grep("h2_z",names(dtmp3))]
dtmp3 <- dtmp3[,-grep("intercept_z",names(dtmp3))]
dtmp3 <- dtmp3[,-grep("Coefficient$",names(dtmp3))]
names(dtmp3) <- gsub("_z","",names(dtmp3))
dtmp3 <- dtmp3[,-grep("_p",names(dtmp3))]
dtmp3 <- dtmp3[,-grep("Prop_SNPs",names(dtmp3))]
dtmp3 <- dtmp3[,-grep("Prop_h2",names(dtmp3))]
dtmp3 <- dtmp3[,-grep("Enrichment",names(dtmp3))]
dtmp3 <- dtmp3[,-grep("base",names(dtmp3))]
dtmp3 <- dtmp3[,-grep("MAFbin",names(dtmp3))]
dtmp3 <- dtmp3[,-grep("CpG_Content",names(dtmp3))]
dtmp3 <- dtmp3[,-grep("GERP.NS",names(dtmp3))]
dtmp3 <- dtmp3[,-grep("MAF_Adj",names(dtmp3))]
dtmp3 <- dtmp3[,-grep("Nucleotide",names(dtmp3))]
dtmp3 <- dtmp3[,-grep("Backgrd",names(dtmp3))]
dtmp3 <- dtmp3[,-grep("Recomb_Rate",names(dtmp3))]
dtmp3 <- dtmp3[,-grep("extend.500",names(dtmp3))]
dtmp3 <- dtmp3[,!(names(dtmp3) %in% c("description","N","N_case","N_control","prevelence","mean_chi2","lambdaGC","intercept","ratio","h2_observed","isNomSig_h2","isNomSig_int","table_desc","h2_liability"))]

dd6 <- melt(dtmp3,c("phenotype","Neff","isBinary","isBonfSig_h2","isBonfSig_int"),variable.name = "annot1",value.name="Coefficient")
dd6$annot <- gsub("..Coefficient","",dd6$annot1)
d6 <- SharedData$new(dd6[dd6$isBonfSig_h2 & dd6$Neff > 100000,])

pp <- plot_ly(d6) %>%
		add_trace(
			  y = ~Coefficient,
			  split = ~annot,
			  type="box",
			  hoverinfo = "x"
	  		) %>%
		layout(showlegend=F,
			   yaxis=list(title="Coefficient Z-score"),
			   margin=list(b=150,l=100), xaxis=list(title="Annotation", tickangle=-30))

bscols(widths=c(10,2),
	config(pp, collaborate = F, showLink=F, displayModeBar=F, displaylogo=F, sendData=F),
	list(
		filter_slider("Neff", "Effective N", d6, ~Neff, step=100, width="100%"),
		filter_select("isBinary","Binary Pheno.", d6, ~isBinary, multiple=T),
		filter_select("isBonfSig_int","Signif. Int.", d6, ~isBonfSig_int, multiple=T)
	)
)
```

<br>

**Note:** For clarity, this plot is limited to phenotypes with significant $h^2$ (after Bonferroni correction) and effective N > 100,000.

</div>

<br>


#### Enrichment

<div class="well">

```{r hist_enrich, echo=FALSE, warnings=FALSE}

dtmp <- dat
dtmp <- dtmp[,-grep("_se",names(dtmp))]
dtmp <- dtmp[,-grep("_z",names(dtmp))]
dtmp <- dtmp[,-grep("_p",names(dtmp))]
dtmp <- dtmp[,-grep("Prop_SNPs",names(dtmp))]
dtmp <- dtmp[,-grep("Prop_h2",names(dtmp))]
dtmp <- dtmp[,-grep("Coefficient",names(dtmp))]
dtmp <- dtmp[,-grep("base",names(dtmp))]
dtmp <- dtmp[,-grep("MAFbin",names(dtmp))]
dtmp <- dtmp[,-grep("CpG_Content",names(dtmp))]
dtmp <- dtmp[,-grep("GERP.NS",names(dtmp))]
dtmp <- dtmp[,-grep("MAF_Adj",names(dtmp))]
dtmp <- dtmp[,-grep("Nucleotide",names(dtmp))]
dtmp <- dtmp[,-grep("Backgrd",names(dtmp))]
dtmp <- dtmp[,-grep("Recomb_Rate",names(dtmp))]
dtmp <- dtmp[,-grep("extend.500",names(dtmp))]
dtmp <- dtmp[,!(names(dtmp) %in% c("description","N","N_case","N_control","prevelence","mean_chi2","lambdaGC","intercept","ratio","h2_observed","isNomSig_h2","isNomSig_int","table_desc","h2_liability"))]

dd2 <- melt(dtmp,c("phenotype","Neff","isBinary","isBonfSig_h2","isBonfSig_int"),variable.name = "annot1",value.name="Enrichment")
dd2$annot <- gsub("..Enrichment","",dd2$annot1)

d4 <- SharedData$new(dd2[dd2$isBonfSig_h2 & dd$Neff > 100000,])
nr <- sum(dd2$isBonfSig_h2 & dd2$Neff > 100000)

pp <- plot_ly(d4, height=600) %>%
		add_trace(
			  y = ~Enrichment,
			  split = ~annot,
			  type="box",
			  hoverinfo = "x"
	  		) %>%
		layout(showlegend=F,
			   margin=list(b=150,l=100),
			   xaxis=list(title="Annotation", tickangle=-30),
			   yaxis=list(zeroline=F)) 

bscols(widths=c(10,2),
	config(pp, collaborate = F, showLink=F, displayModeBar=F, displaylogo=F, sendData=F),
	list(
		filter_slider("Neff", "Effective N", d4, ~Neff, step=100, width="100%"),
		filter_select("isBinary","Binary Pheno.", d4, ~isBinary, multiple=T),
		filter_select("isBonfSig_int","Signif. Int.", d4, ~isBonfSig_int, multiple=T)
	)
)
```

<br>

**Note:** For clarity, this plot is limited to phenotypes with significant $h^2$ (after Bonferroni correction) and effective N > 100,000. The null value for enrichment is 1.

</div>

<br>

### Continuous Annotations {.tabset .tabset-fade}

#### Coefficient Z score

<div class="well">

```{r hist_coef_cont, echo=FALSE}

dtmp2 <- dat
dtmp2 <- dtmp2[,-grep("_se",names(dtmp2))]
dtmp2 <- dtmp2[,-grep("h2_z",names(dtmp2))]
dtmp2 <- dtmp2[,-grep("intercept_z",names(dtmp2))]
dtmp2 <- dtmp2[,-grep("Coefficient$",names(dtmp2))]
names(dtmp2) <- gsub("_z","",names(dtmp2))
dtmp2 <- dtmp2[,-grep("_p",names(dtmp2))]
dtmp2 <- dtmp2[,-grep("Prop_SNPs",names(dtmp2))]
dtmp2 <- dtmp2[,-grep("Prop_h2",names(dtmp2))]
dtmp2 <- dtmp2[,-grep("Enrichment",names(dtmp2))]
dtmp2 <- dtmp2[,-grep("base",names(dtmp2))]
dtmp2 <- dtmp2[,-grep("extend.500",names(dtmp2))]
dtmp2 <- dtmp2[,!(names(dtmp2) %in% levels(dd$annot1))]
dtmp2 <- dtmp2[,!(names(dtmp2) %in% c("description","N","N_case","N_control","prevelence","mean_chi2","lambdaGC","intercept","ratio","h2_observed","isNomSig_h2","isNomSig_int","table_desc","h2_liability"))]

dd3 <- melt(dtmp2,c("phenotype","Neff","isBinary","isBonfSig_h2","isBonfSig_int"),variable.name = "annot1",value.name="Coefficient")
dd3$annot <- gsub("..Coefficient","",dd3$annot1)
d5 <- SharedData$new(dd3[dd3$isBonfSig_h2 & dd3$Neff > 100000,])

pp <- plot_ly(d5) %>%
		add_trace(
			  y = ~Coefficient,
			  split = ~annot,
			  type="box",
			  hoverinfo = "x"
	  		) %>%
		layout(showlegend=F,
			   yaxis=list(title="Coefficient Z-score"),
			   margin=list(b=150,l=100), xaxis=list(title="Annotation", tickangle=-30))

bscols(widths=c(10,2),
	config(pp, collaborate = F, showLink=F, displayModeBar=F, displaylogo=F, sendData=F),
	list(
		filter_slider("Neff", "Effective N", d5, ~Neff, step=100, width="100%"),
		filter_select("isBinary","Binary Pheno.", d5, ~isBinary, multiple=T),
		filter_select("isBonfSig_int","Signif. Int.", d5, ~isBonfSig_int, multiple=T)
	)
)
```

<br>

**Note:** For clarity, this plot is limited to phenotypes with significant $h^2$ (after Bonferroni correction) and effective N > 100,000. Continuous annotations are on differing scales, so their coefficients are not directly comparable.

</div>

<br>

